mouse_microbiome

### 2018-09-06 Manuscript_Mouse_Microbiome.Rmd
major update post review

John Guittar^1,\*^, Aspen Reese^2^, Lawrence David^3^, Timothy Brennhofer^1,4^, Ashley Shade^5,6^, Elena Litchman^1,6,7^

@@

^4^ Grinnell College, 1115 8th Ave, Grinnell, IA 50112, USA

^5^ Department of Microbiology and Molecular Genetics, Department of Plant, Soil and Microbial Sciences, and The Plant Resilience Institute, Michigan State University, East Lansing MI 48840, USA

^6^ Program in Ecology, Evolutionary Biology and Behavior, Michigan State University East Lansing MI 48840, USA

^7^ Department of Integrative Biology, Michigan State University, East Lansing, MI 48824, USA

@@

#rarefy to 5000
tmp <- x[,c(11:ncol(x))]
tmp <- tmp[rowSums(tmp) >= 5000,]
tmp <- rrarefy(tmp, 5000)
x <- bind_cols(x[, c(1:10)], as.data.frame(tmp))

@@

#filter so trait and tax tables have the same taxa.
#create wide version
traits <- filter(traits, otu %in% x$otu)
traits_wide <- spread(traits, trait, val, fill = NA)
tax <- filter(tax, otu %in% x$otu)

@@

p2 <- j %>% 
  filter(!is.na(Pair)) %>%
  mutate(Pair = paste('Cage', Pair)) %>%
  ggplot(aes(x = NMDS1, y = NMDS2, color = Abx, fill = Abx)) + 
  geom_point(aes(alpha = Hour), shape = 21, size = 4) + 
  geom_point(shape = 21, size = 4, fill = NA) + 
  geom_path() + 
  geom_text(aes(label = label), color = 'black', na.rm=TRUE) +
  scale_color_manual(values = c('black','red')) +
  scale_fill_manual(values = c('black','red')) +
  facet_grid(Pair~Abx) + 
  th

@@

j <- x %>%
  group_by(Mouse, Hour) %>%
  mutate(rich = length(unique(otu[abun > 0]))) %>%
  group_by(Mouse, Hour, Abx, Cohoused, Dose, Pair, rich) %>%
  do(shannon = diversity(spread(select(., otu, abun), otu, abun))) %>%
  mutate(shannon_val = shannon[[1]])

#must fix screwy data!!! (when I paste mouse, abx, cohoused, there are some unexpected combinations)
j <- j %>%
  group_by(Mouse, Abx, Cohoused) %>%
  filter(length(shannon_val) > 1) %>%
  ungroup() %>%
  mutate(
    evenness = shannon_val / log(rich),
    Abx = ifelse(Abx == 'Y', 'Abx', 'No abx'),
    Mouse = ifelse(Cohoused == 'Y', paste(Abx, "Cohoused", Mouse), paste(Abx, 'Solitary', Mouse)))

@@

p <- j %>%
  ggplot(aes(x = Hour, y = rich, color = Abx)) +
    geom_point() +
    stat_smooth() +
    scale_color_manual(values = c('red','black')) +
    geom_rect(aes(xmin = 0, xmax = 120, ymin = -Inf, ymax = Inf), alpha = 0.3, fill = 'red', color = NA, data = j %>% group_by(Mouse) %>% filter(row_number() == 1 & Abx == 'Abx')) + 
    facet_wrap(~Mouse, ncol = 3) +
    labs(x = 'Hour', y = 'OTU richness') +
    th

@@

#set the window. The edges are included.
window0 <- 63 
window1 <- 120

j <- x %>%
  group_by(Mouse, Abx, Pair, otu) %>%
  summarise(
    abunPre = if (unique(Abx) == 'Y') sum(abun[Hour <= window0]) else NA,
    abun0 = if (unique(Abx) == 'Y') sum(abun[Hour >= window0 & Hour <= window1]) else sum(abun[Hour < window1]),
    abun1 = sum(abun[Hour > window1]),
    statusAB = if (!is.na(abunPre) & abunPre == 0 & abun0 == 0) 'absent' else 'persistent',
    statusAB = if (!is.na(abunPre) & abunPre > 0 & abun0 == 0) 'extinct' else statusAB,
    ETD = if (statusAB == 'extinct') max(Hour[Hour <= window0 & abun > 0]) else NA,
    ETD = if (statusAB == 'persistent') Inf else ETD,
    status0 = if (abun0 > 0) 'present' else 'absent',
    status1 = if (sum(abun[Hour > window1] > 0)) 'present' else 'absent',
    status = if (status0 == 'absent' & status1 == 'absent') 'absent' else 'persistent',
    status = if (status0 == 'absent' & status1 == 'present') 'immigrant' else status,
    status = if (status0 == 'present' & status1 == 'absent') 'extinct' else status,
    ETA = if (status == 'immigrant') min(Hour[Hour > 97 & abun > 0]) else NA) %>%
  group_by(Pair, otu) %>%
  mutate(
    abun_pair = ifelse(!is.na(Pair) & Abx == 'Y', abun0[Abx == 'N'], NA),
    abun_pair = ifelse(!is.na(Pair) & Abx == 'N', abun0[Abx == 'Y'], abun_pair))

j <- x %>%
  group_by(Mouse, Abx, Pair, otu) %>%
  summarise(
    abun0 = sum(abun[Hour < window0]),
    abun1 = sum(abun[Hour >= window0 & Hour <= window1]),
    abun01 = abun0 + abun1,
    abun2 = sum(abun[Hour > window1]),
    status_treatment = if (abun01 == 0) 'absent' 
                         else if (abun0 > 0 & abun1 == 0) 'extinct' 
                           else 'persistent',
    ETD = if (status_treatment == 'extinct') max(Hour[Hour <= window0 & abun > 0]) 
            else if (status_treatment == 'persistent') Inf 
              else NA,
    status_recovery = if (abun01 == 0 & abun2 == 0) 'absent' 
               else if (abun1 == 0 & abun2 > 0) 'immigrant'
                 else if (abun01 > 0 & abun2 == 0) 'extinct'
                   else 'persistent',
    immigrant_type = if (status_recovery != 'immigrant') 'Not an immigrant'
                       else if (abun01 > 0) 'recovered'
                         else 'new colonist',
    ETA = if (status_recovery == 'immigrant') min(Hour[Hour > 97 & abun > 0]) 
            else NA) %>%
  group_by(Pair, otu) %>%
  mutate(
    abun_pair = ifelse(!is.na(Pair) & Abx == 'Y', abun01[Abx == 'N'], NA),
    abun_pair = ifelse(!is.na(Pair) & Abx == 'N', abun01[Abx == 'Y'], abun_pair))

p <- j %>% 
  filter(status_recovery == 'immigrant' & Abx == 'Y' & abun_pair > 0) %>%
  mutate(time_to_arrival = ETA - 120) %>%
  ggplot(aes(x = abun_pair, y = time_to_arrival)) + 
    geom_jitter(alpha = 0.25, color = 'red', height = 5, width = 0) + 
    scale_x_log10() + 
    stat_smooth(color = 'black', method = 'lm') +
    stat_smooth(color = 'black', lty = 3) +
    labs(x = "OTU relative abundance in cohoused control mouse", y = 'Hours before OTU arrival in Abx treated mouse') +
    th

@@

iwant <- c('Obligate_anaerobe','Sporulation')

@@

>>>>

John Guittar^1,\*^, Aspen Reese^2^, Lawrence David^3^, Ashley Shade^5,6^, Elena Litchman^1,6,7^

@@

^4^ Department of Microbiology and Molecular Genetics, Department of Plant, Soil and Microbial Sciences, and The Plant Resilience Institute, Michigan State University, East Lansing MI 48840, USA

^5^ Program in Ecology, Evolutionary Biology and Behavior, Michigan State University East Lansing MI 48840, USA

^6^ Department of Integrative Biology, Michigan State University, East Lansing, MI 48824, USA

@@

#rarefy to 5000
if (FALSE) {
  tmp <- x[,c(11:ncol(x))]
  tmp <- tmp[rowSums(tmp) >= 5000,]
  tmp <- rrarefy(tmp, 5000)
  x <- bind_cols(x[, c(1:10)], as.data.frame(tmp))
}

@@

#filter so trait and tax tables have the same taxa.
#create wide version
traits <- filter(traits, otu %in% x$otu)
traits_wide <- spread(traits, trait, val, fill = NA)
traits_wide_scaled <- data.frame(otu = traits_wide$otu, sapply(traits_wide[, -1], scale))
traits_wide_scaled <- select(traits_wide_scaled, -Length, -B_vitamins, -Temp_optimum)
tax <- filter(tax, otu %in% x$otu)

@@

p2 <- j %>% 
  filter(!is.na(Pair)) %>%
  mutate(Pair = paste('Cage', Pair)) %>%
  ggplot(aes(x = NMDS1, y = NMDS2, color = Abx, fill = Abx)) + 
  geom_point(aes(alpha = Hour), shape = 21, size = 4) + 
  geom_point(shape = 21, size = 4, fill = NA) + 
  geom_path() + 
  geom_text(aes(label = label), color = 'black', na.rm=TRUE) +
  scale_color_manual(values = c('black','red')) +
  scale_fill_manual(values = c('black','red')) +
  facet_wrap(~Mouse) + 
  th

@@

j <- x %>%
  group_by(Mouse, Hour) %>%
  left_join(tax[, c('otu','Genus','Species')], by = 'otu') %>%
  mutate(rich_spp = length(unique(paste(Genus, Species)[abun > 0 & Species != 'unclassified'])),
         rich = length(unique(otu[abun > 0]))) %>%
  group_by(Mouse, Hour, Abx, Cohoused, Dose, Pair, rich, rich_spp) %>%
  do(shannon = diversity(spread(select(., otu, abun), otu, abun))) %>%
  mutate(shannon_val = shannon[[1]])

j <- j %>%
  ungroup() %>%
  mutate(
    evenness = shannon_val / log(rich),
    Abx = ifelse(Abx == 'Y', 'Abx', 'No abx'),
    Mouse = ifelse(Cohoused == 'Y', paste(Abx, "Cohoused", Mouse), paste(Abx, 'Solitary', Mouse)))

@@

p <- j %>%
  ggplot(aes(x = Hour, y = rich, color = Abx)) +
    geom_point() +
    stat_smooth() +
    scale_color_manual(values = c('red','black')) +
    geom_rect(aes(xmin = 0, xmax = 120, ymin = -Inf, ymax = Inf), alpha = 0.3, fill = 'red', color = NA, data = j %>% group_by(Mouse) %>% filter(row_number() == 1 & Abx == 'Abx')) + 
    geom_hline(yintercept = 0) +
    facet_wrap(~Mouse, ncol = 3) +
    labs(x = 'Hour', y = 'OTU richness') +
    theme_bw()

@@

#set the window. The edges are included.
window0 <- 119 
window1 <- 120

j <- x %>%
  group_by(Mouse, Abx, Pair, otu) %>%
  summarise(
    abun0 = sum(abun[Hour < window0]),
    abun1 = sum(abun[Hour >= window0 & Hour <= window1]),
    abun01 = abun0 + abun1,
    abun2 = sum(abun[Hour > window1]),
    abun_final = if (312 %in% Hour) abun[Hour == 312]
                   else NA,
    status_treatment = if (abun01 == 0) 'absent' 
                         else if (abun0 > 0 & abun1 == 0) 'extinct' 
                           else 'persistent',
    ETD = if (status_treatment == 'extinct') max(Hour[Hour <= window0 & abun > 0]) 
            else if (status_treatment == 'persistent') Inf 
              else NA,
    status_recovery = if (abun01 == 0 & abun2 == 0) 'absent' 
               else if (abun1 == 0 & abun2 > 0) 'immigrant'
                 else if (abun01 > 0 & abun2 == 0) 'extinct'
                   else 'persistent',
    immigrant_type = if (status_recovery != 'immigrant') 'Not an immigrant'
                       else if (abun01 > 0) 'recovered'
                         else 'new colonist',
    ETA = if (status_recovery == 'immigrant') min(Hour[Hour > 97 & abun > 0]) 
            else NA) %>%
  group_by(Pair, otu) %>%
  mutate(
    abun_pair = ifelse(!is.na(Pair) & Abx == 'Y', abun01[Abx == 'N'], NA),
    abun_pair = ifelse(!is.na(Pair) & Abx == 'N', abun01[Abx == 'Y'], abun_pair))

p <- j %>% 
  filter(status_recovery == 'immigrant' & Abx == 'Y') %>%
  mutate(time_to_arrival = ETA - 120) %>%
  ggplot(aes(x = abun_pair + 1, y = time_to_arrival)) + 
    geom_jitter(alpha = 0.25, color = 'red', height = 5, width = 0) + 
    scale_x_log10() + 
    stat_smooth(color = 'black', method = 'lm') +
    stat_smooth(color = 'black', lty = 3) +
    labs(x = "OTU relative abundance in cohoused control mouse + 1", y = 'Hours before OTU arrival in Abx treated mouse') +
    th

@@

iwant <- c('Obligate_anaerobe','Sporulation', 'Copies_16S', 'Gene_number','Motility')

@@

p <- traits %>% 
  left_join(tax, by = 'otu') %>% 
  group_by(trait) %>% 
  arrange(val) %>% 
  mutate(
    rank = seq_along(val), 
    Enterobacteriaceae = Family == 'Enterobacteriaceae') %>% 
  ggplot(aes(x = rank, y = val, color = Enterobacteriaceae)) + 
    geom_point(aes(size = Enterobacteriaceae)) + 
    facet_wrap(~trait, scales = 'free') + 
    scale_color_manual(values = c('black','red')) + 
    th

plot_it(p, name = 'figEnteroTraits', dir = image_dir, width = 6, height = 5)

(ref:figEnteroTraits) Figure `r refs$figEnteroTraits`

`r pagebreak` ![](images\\figEnteroTraits.png)

**(ref:figEnteroTraits):**

j <- x %>%
  group_by(Mouse) %>%
  filter(abun > 0) %>%
  filter(Cohoused != 'dead') %>%
  group_by(Mouse, Hour) %>%
  mutate(rich = length(unique(otu))) %>%
  group_by(Mouse) %>%
  mutate(
    cum_rich = as.numeric(factor(otu, levels = unique(otu))),
    Abx = ifelse(Abx == 'Y', 'Abx', 'No abx'),
    Cohoused = ifelse(Cohoused == 'Y', 'Cohoused', 'Solitary')) %>%
  group_by(Mouse, Hour, Abx, Cohoused, rich) %>%
  summarise(cum_rich = max(cum_rich)) %>%
  group_by(Mouse) %>%
  mutate(new_rich = c(cum_rich[1], cum_rich[2:length(cum_rich)] -
                                   cum_rich[1:length(cum_rich) -1]))

p <- ggplot(j, aes(x = Hour, y = cum_rich, color = Abx)) +
    geom_rect(aes(xmin = 0, xmax = 120, ymin = -Inf, ymax = Inf), alpha = 0.3, fill = 'red', color = NA, data = filter(ungroup(j), row_number() == 1 & Abx == 'Abx')) + 
    geom_vline(xintercept = 120) +
    geom_vline(xintercept = 0) +
    #geom_line(aes(group = Mouse, lty = Cohoused)) +
    geom_point(data = filter(j, cum_rich < 2300)) +
    geom_point(data = filter(j, cum_rich > 2300), shape = 1) +
    stat_smooth(aes(group =  Mouse, lty = Cohoused), se = FALSE,
                data = filter(j, cum_rich < 2300)) +
    scale_color_manual(values = c('red','black')) +
    #facet_wrap(~Abx) +
    expand_limits(y = 0) +
    th +
    labs(x = "Hour", y = 'Cumulative no. of unique OTUs encountered')

plot_it(p, name = 'figCumOTUS', dir = image_dir, width = 6, height = 5)

(ref:figCumOTUS) Figure `r refs$figCumOTUS`

`r pagebreak` ![](images\\figCumOTUS.png)

**(ref:figCumOTUS):**

p <- ggplot(j, aes(x = Hour, y = cum_rich, color = Abx, lty = Cohoused)) +
    geom_point(alpha = 0.2, data = filter(j, cum_rich < 2300)) +
    geom_point(data = filter(j, cum_rich > 2300)) +
    stat_smooth(se = FALSE, data = filter(j, cum_rich < 2300)) + 
    th +
    scale_color_manual(values = c('red','black')) +
    expand_limits(y = 0)

plot_it(p, name = 'figCumOtusByTreat', dir = image_dir, width = 6, height = 5)

(ref:figCumOtusByTreat) Figure `r refs$figCumOtusByTreat`

`r pagebreak` ![](images\\figCumOtusByTreat.png)

**(ref:figCumOtusByTreat):**

p <- j %>% 
  group_by(Mouse, Abx, Cohoused, Hour) %>%
  summarise(rich = unique(rich)) %>%
  ggplot(aes(x = Hour, y = rich, color = Abx, lty = Cohoused)) +
    geom_point(alpha = 0.2) +
    stat_smooth() + 
    scale_color_manual(values = c('red','black')) +
    th +
    labs(y = 'Instantaneous richness')

plot_it(p, name = 'figRichByTreat', dir = image_dir, width = 6, height = 5)

(ref:figRichByTreat) Figure `r refs$figRichByTreat`

`r pagebreak` ![](images\\figRichByTreat.png)

**(ref:figRichByTreat):**

p <- j %>% 
  ggplot(aes(x = Hour, y = new_rich, color = Abx, lty = Cohoused)) +
    geom_point(alpha = 0.2, data = filter(j, new_rich > 0 & (new_rich < 200 | Hour < 200))) +
    geom_point(size = 2, data = filter(j, new_rich > 200 & Hour > 200)) +
    scale_color_manual(values = c('red','black')) +
    stat_smooth(se = FALSE, data = filter(j, new_rich < 200 | Hour < 200)) + 
    th +
    labs(y = 'No. new OTUs encountered')

#j %>% 
#  group_by(Mouse) %>%
#  filter(Hour %in% c(0,23,63,120,240,312)) %>%
#  group_by(Mouse, Abx, Cohoused, Hour) %>%
#  summarise(cum_rich = max(cum_rich)) %>%
#  ggplot(aes(x = factor(Hour), y = cum_rich, color = Abx)) +
#    geom_boxplot(position = 'dodge') +
#    facet_wrap(~Cohoused) +
#    th

plot_it(p, name = 'figNewRichByTreat', dir = image_dir, width = 6, height = 5)

(ref:figNewRichByTreat) Figure `r refs$figNewRichByTreat`

`r pagebreak` ![](images\\figNewRichByTreat.png)

**(ref:figNewRichByTreat):**

j <- x %>% 
  left_join(tax, by = 'otu') %>% 
  filter(Family == 'Enterobacteriaceae') %>%
  group_by(Mouse) %>%
  filter(abun > 0) %>%
  mutate(Cumulative = cummax(as.numeric(factor(otu, levels = unique(otu))))) %>%
  mutate(Abx = factor(ifelse(Abx == 'Y','Abx', 'No abx'), levels = c('No abx', 'Abx'))) %>%
  group_by(Mouse, Abx, Hour) %>%  
  summarise(
    Instantaneous = length(otu), 
    Cumulative = max(Cumulative)) %>%
  gather(var, val, Instantaneous, Cumulative) %>%
  mutate(var = factor(var, unique(var)))

p <- ggplot(j, aes(x = Hour, y = val, color = Abx)) + 
  geom_rect(aes(xmin = 0, xmax = 120, ymin = -Inf, ymax = Inf), alpha = 0.3, fill = 'red', 
            color = NA, data = filter(group_by(j, var), row_number() == 1 & Abx == 'Abx')) + 
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  #geom_line(aes(group = Mouse)) + 
  geom_point(alpha = 0.2) +
  stat_smooth(aes(group = Mouse), se = FALSE) +
  scale_color_manual(values = c('black','red'), name = '') +
  labs(y = "Enterobacteriaceae OTU richness") +
  facet_wrap(~var) +
  th

plot_it(p, name = 'figEntRichHour', dir = image_dir, width = 6, height = 5)

(ref:figEntRichHour) Figure `r refs$figEntRichHour`

`r pagebreak` ![](images\\figEntRichHour.png)

**(ref:figEntRichHour):**

j <- x %>%
  left_join(tax, by = 'otu') %>%
  filter(Abx == 'Y') %>%
  mutate(abun = abun/sum(abun),
         Mouse = paste(Mouse, Cohoused)) %>%
  group_by(Family) %>%
  mutate(fam_abun = sum(abun)) %>%
  ungroup() %>%
  filter(fam_abun > 0.02) %>%
  group_by(Mouse, Family, Hour) %>%
  summarise(abun = sum(abun)) %>%
  ungroup()

p <- ggplot(j, aes(x = Hour, y = abun, color = Family)) +
  geom_rect(aes(xmin = 0, xmax = 120, ymin = -Inf, ymax = Inf), alpha = 0.15, fill = 'red', color = NA, data = j %>% filter(row_number() == 1)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  #facet_wrap(~Mouse) + 
  theme(legend.position = 'bottom',
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.background = element_blank())

plot_it(p, name = 'figFamilyPatterns', dir = image_dir, width = 6, height = 5)

(ref:figFamilyPatterns) Figure `r refs$figFamilyPatterns`

`r pagebreak` ![](images\\figFamilyPatterns.png)

**(ref:figFamilyPatterns):**

j <- x %>%
  filter(Abx == 'Y') %>%
  group_by(Mouse, otu) %>%
  filter(abun[Hour == 0] > 0) %>%
  summarise(abx_response = ifelse(abun[Hour == 120] > 0, 1, 0)) %>%
  ungroup() %>%
  left_join(traits_wide_scaled, by = 'otu') %>%
  select(-Mouse, -otu)

m <- lm(abx_response ~ ., data = j)

tmp <- summary(m)$coef %>%
  as.data.frame() %>%
  do(data.frame(var = row.names(.), .)) %>%
  rename(SE = Std..Error, p.value = Pr...t..) %>%
  filter(var != "(Intercept)") %>%
  mutate(var = factor(var, unique(var[order(t.value)])))

p <- ggplot(tmp, aes(x = var, y = Estimate, color = p.value < 0.05)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_point() +
  labs(x = '', y = 'Mean change in probability of persisting (Y/N) during Abx treatment') +
  coord_flip() +
  th

plot_it(p, name = 'figTraitsDuringAbxBinary', dir = image_dir, width = 6, height = 5)

(ref:figTraitsDuringAbxBinary) Figure `r refs$figTraitsDuringAbxBinary`

`r pagebreak` ![](images\\figTraitsDuringAbxBinary.png)

**(ref:figTraitsDuringAbxBinary):**

j <- x %>%
  filter(Abx == 'Y') %>%
  group_by(Mouse, otu) %>%
  filter(abun[Hour == 0] > 0 | abun[Hour == 120] > 0) %>%
  summarise(delta_rel_abun = abun[Hour == 120] - abun[Hour == 0]) %>%
  ungroup() %>%
  left_join(traits_wide_scaled) %>%
  select(-Mouse, -otu)

m <- lm(delta_rel_abun / 5000 ~ ., data = j)

tmp <- summary(m)$coef %>%
  as.data.frame() %>%
  do(data.frame(var = row.names(.), .)) %>%
  rename(SE = Std..Error, p.value = Pr...t..) %>%
  filter(var != "(Intercept)") %>%
  mutate(var = factor(var, unique(var[order(t.value)])))

p <- ggplot(tmp, aes(x = var, y = Estimate, color = p.value < 0.05)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_point() +
  labs(y = 'Mean Change in rel. abun. over Abx treatment', x = '') +
  coord_flip() +
  th

plot_it(p, name = 'figTraitsDuringAbxContinuous', dir = image_dir, width = 6, height = 5)

(ref:figTraitsDuringAbxContinuous) Figure `r refs$figTraitsDuringAbxContinuous`

`r pagebreak` ![](images\\figTraitsDuringAbxContinuous.png)

**(ref:figTraitsDuringAbxContinuous):**

j <- x %>%
  filter(!is.na(Pair)) %>%
  group_by(Pair, otu) %>%
  mutate(Control_abun = sum(abun[Abx == 'N']) / length(unique(Hour))) %>%
  filter(Abx == 'Y') %>%
  group_by(Mouse, otu, Control_abun) %>%
  filter(abun[Hour == 120] > 0) %>%
  summarise(abx_response = ifelse(abun[Hour == max(Hour)] > 0, 1, 0)) %>%
  ungroup() %>%
  left_join(traits_wide_scaled, by = 'otu') %>%
  mutate(Control_abun = scale(Control_abun)) %>%
  select(-Mouse, -otu)

m <- lm(abx_response ~ ., data = j)

tmp <- summary(m)$coef %>%
  as.data.frame() %>%
  do(data.frame(var = row.names(.), .)) %>%
  rename(SE = Std..Error, p.value = Pr...t..) %>%
  filter(var != "(Intercept)") %>%
  mutate(var = factor(var, unique(var[order(Estimate)])))

p <- ggplot(tmp, aes(x = var, y = Estimate, color = p.value < 0.05)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_point() +
  labs(x = '', y = 'Mean change in probability of persisting (Y/N) over Abx recovery') +
  coord_flip() +
  th

plot_it(p, name = 'figTraitsAfterAbxBinary', dir = image_dir, width = 6, height = 5)

(ref:figTraitsAfterAbxBinary) Figure `r refs$figTraitsAfterAbxBinary`

`r pagebreak` ![](images\\figTraitsAfterAbxBinary.png)

**(ref:figTraitsAfterAbxBinary):**

j <- x %>%
  filter(!is.na(Pair)) %>%
  group_by(Pair, otu) %>%
  mutate(Control_abun = sum(abun[Abx == 'N']) / length(unique(Hour))) %>%
  filter(Abx == 'Y') %>%
  group_by(Mouse, otu, Control_abun) %>%
  filter(abun[Hour == 120] > 0) %>%
  summarise(delta_rel_abun = abun[Hour == 120] - abun[Hour == max(Hour)]) %>%
  ungroup() %>%
  left_join(traits_wide_scaled, by = 'otu') %>%
  mutate(Control_abun = scale(Control_abun)) %>%
  select(-Mouse, -otu)

m <- lm(delta_rel_abun ~ ., data = j)

tmp <- summary(m)$coef %>%
  as.data.frame() %>%
  do(data.frame(var = row.names(.), .)) %>%
  rename(SE = Std..Error, p.value = Pr...t..) %>%
  filter(var != "(Intercept)") %>%
  mutate(var = factor(var, unique(var[order(Estimate)])))

p <- ggplot(tmp, aes(x = var, y = Estimate, color = p.value < 0.05)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_point() +
  labs(y = 'Mean Change in rel. abun. over Abx recovery', x = '') +
  coord_flip() +
  th

plot_it(p, name = 'figTraitsAfterAbxContinuous', dir = image_dir, width = 6, height = 5)

(ref:figTraitsAfterAbxContinuous) Figure `r refs$figTraitsAfterAbxContinuous`

`r pagebreak` ![](images\\figTraitsAfterAbxContinuous.png)

**(ref:figTraitsAfterAbxContinuous):**

j <- x %>%
  filter(!is.na(Pair)) %>%
  group_by(Pair, otu) %>%
  mutate(Control_abun = sum(abun[Abx == 'N']) / length(unique(Hour))) %>%
  group_by(Mouse, otu, Control_abun) %>%
  summarise(
    Initial_abun = abun[Hour == 0],
    End_abx_abun = abun[Hour == 120],
    final_abun = abun[Hour == max(Hour)],
    Persistence = ifelse(abun[Hour == 0] > 0 & abun[Hour == 120] > 0, 1, 0),
    Persistence = ifelse(sum(abun[Hour %in% c(0,120)]) == 0, NA, Persistence)) %>%
  ungroup() %>%
  mutate(Control_abun = scale(log10(Control_abun+1)),
         Initial_abun = scale(log10(Initial_abun+1)),
         End_abx_bun = scale(log10(End_abx_abun+1)),
         Persistence = scale(Persistence)) %>%
  select(-Mouse, -otu)

m <- lm(final_abun ~ ., data = j)

tmp <- summary(m)$coef %>%
  as.data.frame() %>%
  do(data.frame(var = row.names(.), .)) %>%
  rename(SE = Std..Error, p.value = Pr...t..) %>%
  filter(var != "(Intercept)") %>%
  mutate(var = factor(var, unique(var[order(Estimate)])))

p <- ggplot(tmp, aes(x = var, y = Estimate, color = p.value < 0.05)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_point() +
  labs(x = '', y = 'Predictors of post-recovery rel. abundance') +
  coord_flip() +
  th

plot_it(p, name = 'figAbxPlusRecoveryPredictors', dir = image_dir, width = 6, height = 5)

(ref:figAbxPlusRecoveryPredictors) Figure `r refs$figAbxPlusRecoveryPredictors`

`r pagebreak` ![](images\\figAbxPlusRecoveryPredictors.png)

**(ref:figAbxPlusRecoveryPredictors):**

j <- x %>%
  group_by(Mouse) %>%
  filter(abun > 0) %>%
  filter(Cohoused != 'dead') %>%
  group_by(Mouse) %>%
  mutate(
    cum_rich = as.numeric(factor(otu, levels = unique(otu))),
    Abx = ifelse(Abx == 'Y', 'Abx', 'No abx'),
    Cohoused = ifelse(Cohoused == 'Y', 'Cohoused', 'Solitary')) %>%
  group_by(Mouse, Hour, Abx, Cohoused) %>%
  summarise(cum_rich = max(cum_rich))

p <- ggplot(j, aes(x = Hour, y = cum_rich, color = Abx)) +
    geom_rect(aes(xmin = 0, xmax = 120, ymin = -Inf, ymax = Inf), alpha = 0.3, fill = 'red', color = NA, data = filter(ungroup(j), row_number() == 1 & Abx == 'Abx')) + 
    geom_vline(xintercept = 120) +
    geom_vline(xintercept = 0) +
    #geom_line(aes(group = Mouse, lty = Cohoused)) +
    geom_point(data = filter(j, cum_rich < 2300)) +
    geom_point(data = filter(j, cum_rich > 2300), shape = 1) +
    stat_smooth(aes(group = Mouse, lty = Cohoused), se = FALSE, data = filter(j, cum_rich < 2300)) +
    scale_color_manual(values = c('red','black')) +
    facet_wrap(~Abx) +
    expand_limits(y = 0) +
    th +
    labs(x = "Hour", y = 'Cumulative no. of unique OTUs encountered')

plot_it(p, name = 'figCumulativeOTUs', dir = image_dir, width = 6, height = 5)

(ref:figCumulativeOTUs) Figure `r refs$figCumulativeOTUs`

`r pagebreak` ![](images\\figCumulativeOTUs.png)

**(ref:figCumulativeOTUs):**
